### 2022-03-04 unif/unif.tex
examples

\begin{lemma}[$\jp L{3\text{ (1)}}$]\label{lem:principal-lemma}Let $\varrho$ be a substitution
    and $X$, $Y$ be free variables such that $\varrho\left(X\,\seq s\right)=\varrho\left(Y\,\seq t\right)$
    for some term tuples $\seq s$ and $\seq t$. Then for every opponent
    pair $u$, $v$ in $\varrho X$ and $\varrho Y$
    (Definition \ref{unif:def:opponents}), the head of $u$ or $v$ is a parameter
    of $\varrho X$ or $\varrho Y$.
    \end{lemma}
  %

@@

  \begin{defi}
  [$\jp D2$]\label{unif:def:omega-simple}An occurrence of a parameter $x$ of term $t$
  in the body of $t$ is $\omega$\emph{-simple }if both
  \begin{enumerate}
  \item the arguments of $x$ are distinct and are exactly (the $\eta$-long
  forms of) all of the variables bound in the body of $t$, and
  \item this occurrence of $x$ is not in an argument of any parameter
  of $t$.
  This definition is slightly too restrictive for our purposes.
  It is unfortunate that condition~1 requires $x$ to be applied to
  \emph{all} instead of just some of the bound variables. The JP
  proof would probably work with such a relaxation, and the definition
  would then cover all cases where eager decomposition is possible. 
  However, to reuse the JP lemmas, we stick to the original
  notion of $\omega$-simplicity and introduce the following relaxation:
  \begin{defi}
  An occurrence of a parameter $x$ of term $t$ in the body of $t$
  is \emph{base-simple} if it is $\omega$-simple or both
  \begin{enumerate}
  \item $x$ is of base type, and
  \item this occurrence of $x$ is not in an argument of any parameter
  of $t$.

>>>>

\begin{exa} To determine the opponent pairs of $\lamx{\cst{g} \, (\cst{g} x)}$
  and $\lamx{\cst{g} \, x}$ we first $\alpha$-rename the second term to 
  $\lam{y}{\cst{g}\,{y}}$. Then, we identify the common context of $\cst{g} \, (\cst{g} x)$ and $\cst{g} \, y$
  as $\cst{g} \, \square $, which leads us to the opponent pair $(\cst{g} \, x, y)$.
\end{exa}

\begin{lemma}[$\jp L{3\text{ (1)}}$]\label{lem:principal-lemma}Let $\varrho$ be a substitution
    and $X$, $Y$ be free variables such that $\varrho\left(X\,\seq s\right)=\varrho\left(Y\,\seq t\right)$
    for some term tuples $\seq s$ and $\seq t$. Then for every opponent
    pair $u$, $v$ in $\varrho X$ and $\varrho Y$
    (Definition \ref{unif:def:opponents}), the head of $u$ or $v$ is a parameter
    of $\varrho X$ or $\varrho Y$.
    \end{lemma}
  %
\begin{exa}
  Let $\varrho$ be the substitution $\{ X \mapsto \lamx{\cst{f} \, x}, Y
  \mapsto \lamx{x} \}$. Clearly, $\varrho$ unifies the problem $X \, \cst{a}
  \unif Y \, (\cst{f} \, \cst{a})$. The opponent pair for $\varrho\,X$ and $\varrho\,Y$
  is $(\cst{f} \, x, y)$ which clearly adheres to the conclusion of Lemma \ref{lem:principal-lemma}.
%
  Intuitively, one of the consequences of this lemma is that any solution for
  flex-flex problems that does not mention parameters will not have opponent pairs. 
  For example, one solution for the 
  problem $X \, \cst{a}
  \unif Y \, (\cst{f} \, \cst{a})$ is $\varrho = \{ X \mapsto \cst{c}, Y \mapsto \cst{c} \}$
  which has no opponent pairs.
\end{exa}

@@

  \begin{defi}
  [$\jp D2$]\label{unif:def:omega-simple}An occurrence of a parameter $x$ of term $t$
  in the body of $t$ is $\omega$\emph{-simple }if both
  \begin{enumerate}
  \item the arguments of $x$ are distinct and are exactly (the $\eta$-long
  forms of) all of the variables bound in the body of $t$, and
  \item this occurrence of $x$ is not in an argument of any parameter
  of $t$.
  \end{enumerate}
  \end{defi}

  \begin{exa}
    % \newcommand{\boldboundvar}[1]{\ensuremath{\mathit{\mathbf{#1}}}}
    Following Example 1 in JP's work \cite{jp-76-unif}, we denote
    $\omega$-simple occurrences of parameters using boldface. For example,
    $\lambda xy. \, \cst{f} \, \mathbfit{x} \, \mathbfit{y}$ and $\lamx{\cst{g}
    \, (\lam{y}{\mathbfit{x}\, y})}$ have several $\omega$-simple occurrences.
    In the following two examples, boldface occurrences are \emph{not}
    $\omega$-simple: $\lamx{\cst{g} \, (\lam{y}{\cst{f} \, y \, \mathbfit{x}})},\;
    \lam{xy}{\cst{h} \, (x \, \mathbfit{y})}$. In the first case $x$ is not applied to parameter $y$ and in the second $y$
    occurs as an argument to $x$.
  \end{exa}
  This definition is slightly too restrictive for our purposes.
  It is unfortunate that condition~1 requires $x$ to be applied to
  \emph{all} instead of just some of the bound variables. The JP
  proof would probably work with such a relaxation, and the definition
  would then cover all cases where eager decomposition is possible. 
  However, to reuse the JP lemmas, we stick to the original
  notion of $\omega$-simplicity and introduce the following relaxation:
  \begin{defi}
  An occurrence of a parameter $x$ of term $t$ in the body of $t$
  is \emph{base-simple} if it is $\omega$-simple or both
  \begin{enumerate}
  \item $x$ is of base type, and
  \item this occurrence of $x$ is not in an argument of any parameter
  of $t$.
  \end{enumerate}
  \end{defi}
